---
title: "Applied Bayesian Modeling - module 6"
author: "Leontine Alkema"
date: "September 17, 2022"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(brms)
library(ggmcmc) # used to make density plots for priors and posteriors 

```

# Radon data 

Read in the radon data and process (copied from earlier module)
```{r}
# house level data
d <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/srrs2.dat"), 
                header=T, sep=",")

# deal with zeros, select what we want, make a fips (county) variable to match on 
d <- d %>% 
  mutate(activity = ifelse(activity==0, 0.1, activity)) %>% 
  mutate(fips = stfips * 1000 + cntyfips) %>%   
  dplyr::select(fips, state, county, floor, activity)

# county level data
cty <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/cty.dat"), 
                  header = T, sep = ",")
cty <- 
  cty %>% 
  mutate(fips = 1000 * stfips + ctfips) %>% 
  dplyr::select(fips, Uppm) %>%
  rename(ura_county = (Uppm))

dmn <- d %>% 
  filter(state=="MN") %>% # Minnesota data only
  dplyr::select(fips, county, floor, activity) %>% 
  left_join(cty) 

y <- log(dmn$activity)
ybar <- mean(y)
sd.y <- sd(y)
n <- length(y)

```


# Model fitting using lm and brms

data 
```{r}
dat <- dmn %>%
  mutate(y = log(activity))
```


## frequentist/tradtional

simple fit 
```{r}
fit_lm <-  lm(y ~ 1, data = dat) 
summary(fit_lm)
```

here the same model fitting, just coded up differently in a pipe, with output in nice form
```{r}
dat %>%
  lm(y ~ 1, data = .) %>%
  broom::tidy(conf.int = TRUE, conf.level = 0.95) %>%
  select(-statistic, -p.value) #%>%
#  knitr::kable(format = "latex", digits = 2)

```


## Bayesian regression 
```{r, message  = FALSE}
fit <- brm(y ~ 1, data = dat,
        chains = 4, iter = 1000, warmup = 500, cores = getOption("mc.cores", 4))
```


Quick summary overview
```{r}
summary(fit) 
```

Aside on summaries: you can pull out whatever you're interested in
```{r}
posterior_summary(fit, probs = c(.025, .25, .75, .975), pars = "b_Intercept")
```


Note that the object fit now contains lots of info, we will go through some here 
```{r}
names(fit)
```

# Plots to show posterior samples

Joint density, using bins
```{r}
posterior_samples(fit, pars = c("b_Intercept", "sigma"))%>%
  ggplot(aes(x = sigma, y = b_Intercept)) +
  geom_bin2d() +
  theme_bw()

```

Joint density, estimated
```{r}
posterior_samples(fit, pars = c("b_Intercept", "sigma"))%>%
  ggplot(aes(x = sigma, y = b_Intercept)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")+
  theme_bw()

```


marginal densities
```{r}
posterior_samples(fit, pars = c("b_Intercept"))%>%
  ggplot(aes(x = b_Intercept)) +
  geom_histogram() +
  theme_bw()
```

```{r}
posterior_samples(fit, pars = c("sigma"))%>%
  ggplot(aes(x = sigma)) +
  geom_histogram() +
  theme_bw()
```



# MCMC diagnostics 
Traceplots and posterior densities. Note that you can find the help function with 
```{r}
#?plot.brmsfit
```

```{r}
plot(fit, variable = c("b_Intercept", "sigma")) 
```

Check Rhat and effective sample size
```{r}
summary(fit) 
#names(summary(fit))
summary(fit)$fixed
summary(fit)$spec_pars
```



# MCMC diagnostics in a less ideal setting...

Please note: These fits are based on settings that are NOT recommended. We are just creating an example here of a fit where the MCMC diagnostics (rightly) show that there are issues with the sampling. 
```{r}
fit_bad <- brm(y ~ 1, data = dat,
        chains = 4, iter = 400, cores = getOption("mc.cores", 4),
        control = list(adapt_delta = 0.4, max_treedepth = 4)
        # these are NOT recommended options, trying to create problems here!
)
```


```{r}
plot(fit_bad,  variable = c("b_Intercept",  "sigma")) 
```
```{r}
summary(fit_bad)
```

We see that the effective sample size are too low. Rhat is no longer equal to 1 (although still less than 1.05). 


## comparison fit

Here we create a fit with the same number of iterations for comparison 
```{r}
fit2_short <- brm(y ~ 1, data = dat,
        chains = 4, iter = 400, cores = getOption("mc.cores", 4))
```


```{r}
plot(fit2_short,  variable = c("b_Intercept",  "sigma")) 
```

```{r}
summary(fit2_short)
```


Much larger effective sample sizes! 