---
title: "Applied Bayesian modeling - HW3"
header-includes:
    - \usepackage{bm}
output:
  pdf_document: default
  html_document:
    df_print: paged
---


Score: The maximum number of points in this HW is 20 points, with 3 points extra credit. This HW counts for 1.5 HWs, so points are rescaled such that a "full pass" score is 150%. Specifically, for calculating your final HW score, the points will be rescaled to a maximum score of (20+3)/20*150% = 172.5%. 




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
```

# Exercise 1: Fit a Bayesian model using brm and check and interpret the output (4 points)

We continue with IQ data, as introduced in HW2. For this HW, data set iq_scores.csv contains 10 iq-scores, sampled from a town in let's say, your favorite country. 

Use brm to fit the following model to the IQ data:
\begin{eqnarray*}
y_i|\theta_i, \sigma^2 &\sim& N(\theta_i, \sigma^2) \text{(independent), for } i = 1, 2, \hdots, n; \\
\mu &\sim& N(100, 15^2);\\
\sigma &\sim& \text{use the brm-default}.
\end{eqnarray*} 

Then answer the following questions:  
(i) Plot a histogram of the posterior samples of $\mu$ and report a posterior point estimate and 80% CI for $\mu$.  
(ii) Plot a histogram of the posterior samples of $\sigma$ and report a posterior point estimate and 80% CI for $\sigma$.  
(iii) Can you report a posterior point estimate and 80% CI for $\mu/\sigma$? If yes, do so. If not, why not?   

Note that the prior for $\mu$ can be specified with an additional argument in brm (as illustrated in optional material in module 5), as follows:
```{r}
#mu_prior <- set_prior("normal(100,15)", class = "Intercept")
#fit  <- brm(y ~ 1, prior = c(mu_prior), 
#           your other usual arguments)
```


# Exercise 2: Compare and contrast the MCMC diagnostics of two different model fits (4 points)

Continue with the IQ data from Q1 (with y= IQ scores)  to fit the model  as specified below. Present and briefly summarize resulting MCMC diagnostics for $\mu$ (traceplots, Rhat, effective sample sizes). Then and comment on whether this model fit can be used for summarizing information regarding $\mu$. If not, why not? 


```{r}
#fit_bad <- brm(y ~ 1, data = dat,
#        chains = 4, iter = 200, cores = getOption("mc.cores", 4),
#        control = list(adapt_delta = 0.6, max_treedepth = 4)
#         # these are NOT recommended options, trying to create problems here!
#        )
```

# Exercise 3: see separate pdf (8 points)

# Exercise 4 (4 points)
Continue with the marriage data set from exercise 3.  
The goal in this exercise is to predict the age at first marriage for a randomly sampled woman in an ethnic group for which we have not yet observed any data, using the model and data from exercise 3.

(a) Obtain samples from the predictive posterior density for the age at first marriage and visualize the samples in a histogram. In your answer, include R code (that does NOT use the predict function from brms) as well as a write up in equations how you obtained the samples. Make sure to introduce notation first to explain what you're sampling. 

(b) Use the samples to construct a point prediction and 95% prediction interval for age at first marriage. In your answer, include the expression used for calculating the point prediction from the samples. 

(c) What is the probability that the observed age at first marriage will be greater than $\bar{y}$? In your answer, include the expression used for calculating this probability from the samples. 


# Exercise 5 (extra credit, 3 points)

Obtain the joint distribution of $(y_1, y_2)|\mu_{\alpha}, \sigma_{\alpha}^2, \sigma_{y}^2$ in the setting where $j[1] \neq j[2]$ and in the setting where $j[1] = j[2]$.  

Hints:  

- First consider the univariate distribution of $y_i|\mu_{\alpha}, \sigma_{\alpha}^2, \sigma_{y}^2$.  
- You may find it helpful to write $y_i$ as the sum of parameters (thus random variables) $\mu_{\alpha}$, a parameter to capture the variability across counties, and a parameter to capture variability across observations.






