---
title: "Applied Bayesian modeling - HW3, V1"
header-includes:
    - \usepackage{bm}
output:
  pdf_document: default
  html_document:
    df_print: paged
---

I am posting the HW-V1 to have questions to work on in class on 9/22/after watching module 6. Note that there may be additional questions added later.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
```

# Exercise 1: Fit a Bayesian model using brm and check and interpret the output

We continue with IQ data, as introduced in HW2. For this HW, data set iq_scores.csv contains 10 iq-scores, sampled from a town in let's say, your favorite country. 

Use brm to fit the following model to the IQ data:
\begin{eqnarray*}
y_i|\theta_i, \sigma^2 &\sim& N(\theta_i, \sigma^2) \text{(independent), for } i = 1, 2, \hdots, n; \\
\mu &\sim& N(100, 15^2);\\
\sigma &\sim& \text{use the brm-default}.
\end{eqnarray*} 

Then answer the following questions:  
(i) Plot a histogram of the posterior samples of $\mu$ and report a posterior point estimate and 80% CI for $\mu$.  
(ii) Plot a histogram of the posterior samples of $\sigma$ and report a posterior point estimate and 80% CI for $\sigma$.  
(iii) Can you report a posterior point estimate and 80% CI for $\mu/\sigma$? If yes, do so. If not, why not?   

Note that the prior for $\mu$ can be specified with an additional argument in brm (as illustrated in optional material in module 5), as follows:
```{r}
#mu_prior <- set_prior("normal(100,15)", class = "Intercept")
#fit  <- brm(y ~ 1, prior = c(mu_prior), 
#           your other usual arguments)
```


## Exercise 2: Compare and contrast the MCMC diagnostics of two different model fits

Continue with the IQ data from Q1 (with y= IQ scores)  to fit the model  as specified below. Present and briefly summarize resulting MCMC diagnostics for $\mu$ (traceplots, Rhat, effective sample sizes). Then and comment on whether this model fit can be used for summarizing information regarding $\mu$. If not, why not? 


```{r}
#fit_bad <- brm(y ~ 1, data = dat,
#        chains = 4, iter = 200, cores = getOption("mc.cores", 4),
#        control = list(adapt_delta = 0.6, max_treedepth = 4)
#         # these are NOT recommended options, trying to create problems here!
#        )
```


