---
title: "Applied Bayesian Modeling - module 5"
author: "Leontine Alkema"
date: "August 29, 2022"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```


# Illustration of MC approximation

Note: make sure to set a seed to make results that involve sampling reproducible
```{r}
set.seed(1234)
samp <- rnorm(10000)
mean(samp)


tibble(samp) %>%
  ggplot() +
  geom_histogram(aes(x = samp, y =..density..), fill = "green", alpha = 0.5) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1)) +
 ggtitle("10K samples")

```

# Fit a Bayesian model using brms

In this lab, we will first install the brms and rstan packages for Bayesian model fitting. The goal is to make sure that everyone has a working set-up, to not be held back by that next week when we're starting to get into details of brms/rstan.  

## Installation
Steps are to:  
- install a C++ compiler (needed for Stan to run), see here   
<https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started#configuring-c-toolchain>  
- followed by installing the rstan and brms packages, see  
<https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started#latest-development-version-226x---unreleased>  
<https://github.com/paul-buerkner/brms#how-do-i-install-brms>

## Verfiy that your rstan installation is working alright

You can then verify your rstan installation, see here 
<https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started#verifying-installation>
I.e., run the following 
```{r}
#example(stan_model, package = "rstan", run.dontrun = TRUE)
```

## Fit a model using brms

Let's fit some models to the radon data 

```{r}
library(brms)
```

Read in the radon data and process (copied from earlier module)
```{r}
# house level data
d <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/srrs2.dat"), 
                header=T, sep=",")

# deal with zeros, select what we want, make a fips (county) variable to match on 
d <- d %>% 
  mutate(activity = ifelse(activity==0, 0.1, activity)) %>% 
  mutate(fips = stfips * 1000 + cntyfips) %>%   
  dplyr::select(fips, state, county, floor, activity)

# county level data
cty <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/cty.dat"), 
                  header = T, sep = ",")
cty <- 
  cty %>% 
  mutate(fips = 1000 * stfips + ctfips) %>% 
  dplyr::select(fips, Uppm) %>%
  rename(ura_county = (Uppm))

dmn <- d %>% 
  filter(state=="MN") %>% # Minnesota data only
  dplyr::select(fips, county, floor, activity) %>% 
  left_join(cty) 

y <- log(dmn$activity)
ybar <- mean(y)
sd.y <- sd(y)
n <- length(y)

```

Just fit one model to check that brms is working ok. Here we are again estimating the mean $\mu$, just like in module 3, using brms-default priors for $\mu$ and $\sigma$ (to be discussed further). To estimate the mean, we can use a regression model (and corresponding formula, like in an  lm call) with an intercept only. 

Note: I save the fitted model in a directory "output", such that when knitting, the model is not rerun (see argument of brms, if you're interested). It's a handy option, just watch out with it when you're still updating your data inputs (as it won't update!). 
```{r}
#dir.create("output") # if you don't have an output directory yet 
fit <- brm(y ~ 1, family = gaussian(), data = tibble(y), 
            cores = getOption("mc.cores", 4),
           file = "output/mod5_fit" # adding this to save the fit, 
           # then when knitting, the model is not rerun (see arguments)
            )
```

Check the output
```{r}
fit
```



## Optional extra fits

Fit models with different combinations of  
- Mu: informative vs vague prior for mu  
- sigma fixed at sd(y) or estimated (with default vague prior assigned by brms)

Specifying prior settings for mu (we'll discuss details later re how this is coded):
```{r}
mu_informativepriorhw1 <- set_prior("normal(0,0.1)", class = "Intercept")
mu_vagueprior <- set_prior("normal(0,100)", class = "Intercept")
sigma_fixed <- set_prior("constant(0.86)", class = "sigma")
```

Now fit
```{r}
fit_muinformative_sigmafixed  <- brm(y ~ 1, family = gaussian(), data = tibble(y), 
            cores = getOption("mc.cores", 4),
            prior = c(mu_informativepriorhw1, sigma_fixed),
             file = "output/mod5_fit_muinformative_sigmafixed"
            )
```


```{r}
fit_muvague_sigmafixed  <- brm(y ~ 1, family = gaussian(), data = tibble(y), 
            cores = getOption("mc.cores", 4),
            prior = c(mu_vagueprior, sigma_fixed),
            file = "output/mod5_fit_muvague_sigmafixed"
            )
```


```{r}
fit_muvague_estimatesigma  <- brm(y ~ 1, family = gaussian(), data = tibble(y), 
            cores = getOption("mc.cores", 4),
            prior = c(mu_vagueprior),
            file = "output/mod5_fit_muvague_estimatesigma"  
            )
```


```{r}
fit_muinformative_estimatesigma  <- brm(y ~ 1, family = gaussian(), data = tibble(y), 
            cores = getOption("mc.cores", 4),
            prior = c(mu_informativepriorhw1),
            file = "output/mod5_fit_muinformative_estimatesigma"
            )
```





```{r}
fit_muinformative_sigmafixed
```

```{r}
fit_muvague_sigmafixed
```

```{r}
fit_muinformative_estimatesigma
```

```{r}
fit_muvague_estimatesigma
```




