---
title: "Applied Bayesian Modeling - module 7"
author: "Leontine Alkema"
date: "September 25, 2022"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(brms)
```

# Read in radon data 
Read in the radon data and process (copied from earlier module)
```{r}
# house level data
d <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/srrs2.dat"), 
                header=T, sep=",")

# deal with zeros, select what we want, make a fips (county) variable to match on 
d <- d %>% 
  mutate(activity = ifelse(activity==0, 0.1, activity)) %>% 
  mutate(fips = stfips * 1000 + cntyfips) %>%   
  dplyr::select(fips, state, county, floor, activity)

# county level data
cty <- read.table(url("http://www.stat.columbia.edu/~gelman/arm/examples/radon/cty.dat"), 
                  header = T, sep = ",")
cty <- 
  cty %>% 
  mutate(fips = 1000 * stfips + ctfips) %>% 
  dplyr::select(fips, Uppm) %>%
  rename(ura_county = (Uppm))

dmn <- d %>% 
  filter(state=="MN") %>% # Minnesota data only
  dplyr::select(fips, county, floor, activity) %>% 
  left_join(cty) 

```



## More data processing for multilevel modeling

Some more data processing first, to produce a data set that has county info and to produce the plot with means from the slides. In the data set:

- $y_i$ is log(activity)
- county gives county name (fips gives the unique county ID)
- $x_i$ is floor
- $u_i$ is log_ur = log(ura_county)

(the last two are added for module 8, when including predictors)

```{r}
dat <- 
  dmn%>%
  mutate(y = log(activity), log_ur = log(ura_county))
head(dat)
```

Create summary data set with info for each county:
```{r}
# to plot observations and county means ~ sample sizes, 
# easier to see if sample sizes are slighly jittered
set.seed(12345)

datcounty <- dat %>%
  group_by(fips) %>%
  summarize(nhouses = n(), ybar = mean(y), county = county[1], log_ur = log_ur[1]) %>%
  mutate(nhouses_jitter = nhouses*exp(runif (length(nhouses), -.1, .1)))
ngroups <- dim(datcounty)[1]
head(datcounty)
```


```{r}
ybarbar <- mean(dat$y) # population (here state) mean
```

```{r}
datcounty %>% 
  ggplot(aes(x = nhouses_jitter, y = ybar)) +
  geom_point() + 
  geom_hline(mapping = aes(yintercept = ybarbar)) + 
  theme_bw()
```

# Model fitting

```{r}
fit <- brm(y ~ (1|county), 
        data = dat, 
        iter = 1000,
        chains = 4,
        cores = getOption("mc.cores", 4))
```

Summary of model fit: 
```{r}
summary(fit)
```

# Visualizing the group-level mean parameters 

Coefficients can be obtained using coef(fit), you can get the help file here:
```{r}
#?coef.brmsfit 
```

Just showing some function calls here first, ie for  mu_alpha:
```{r}
fixef(fit)
```


eta = alpha - mu_alpha (as compared to notation in slides), labeled here as random effects 
```{r}
eta <- as_tibble(ranef(fit)$county[,,"Intercept"], rownames = "county")
head(eta)
```

To get the alpha = eta + mu_alpha, we can use the following call
```{r}
alphas <-
  coef(fit, summary = T)$county %>% 
  as_tibble(rownames = "county") %>%
  rename(alph = Estimate.Intercept)
alphas 
```

Make the plot of alpha ~ ybar
```{r}
alphas %>%
  left_join(datcounty) %>%
  ggplot(aes(y = alph, x = ybar, size = nhouses)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)
```

Plot of alpha - ybar
```{r}
alphas %>%
  left_join(datcounty) %>%
  ggplot(aes(y = alph - ybar, x = nhouses)) +
  geom_point() +
  geom_hline(yintercept = 0)
```

